Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)

### 2017-02-15 csf-2017/paper.tex
first pass related work

Capability machines originate with Dennis and Van
Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to
\citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview of
previous work in the area. The capability machine formalised in
Section~\ref{sec:capab-mach-with} is a simple but representative model, modeled
mainly after the M-Machine~\cite{Carter:1994:HSF:195473.195579} (the enter
pointers resemble the M-Machine's) and
CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the memory
capabilities resemble CHERI's). The latter is a recent and relatively mature
capability machine, which combines capabilities with a virtual memory approach,
in the interest of backwards compatibility and gradual adoption. Our local
capabilities are inspired by CHERI's, although as discussed, they can be passed
accross module boundaries, contrary to what is enforced by CHERI's default CCall
implementation.

There are plenty of other papers that enforce well-bracketed control flow at a
low level, but most are restricted to preventing particular types of attacks and
enforce only partial correctness of control flow. This includes particularly the
line of work on \emph{control flow integrity}~\citep{abadi_control-flow_2005}.
This work uses a quite different attacker model than us: they assume an attacker
that is not able to execute code, but can overwrite arbitrary data at any time
during execution (as a model for buffer overflows). By checking the address of
every indirect jump and using memory access control to prevent overwriting code,
this work enforces what they call control flow integrity, formalised as the
property that every jump will follow a legal path in the control flow graph.
As discussed in Section~\ref{sec:discussion}, this correctness property does not
take into account temporal properties and seems hard to use for reasoning.

@@

Our logical relation is a unary, step-indexed Kripke logical relation (SKLR)
with recursive
worlds~\cite{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
capability safety in a high-level JavaScript-like lambda calculus. Our
fundamental theorem is similar to theirs and can be seen as a formulation of
capability safety of the capability machine. Because we are not interested in
externally observable side-effects (like console output or memory access
traces), we do not require their notion of effect parametricity. Our logical
relation uses several ideas from previous work, like Kripke worlds with regions
containing state transition systems~\citep{Ahmed:popl09}, public/private future
worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion),
biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.

El-Korashy also defined a formal model of a capability machine, namely CHERI,
and uses it to prove a compartmentalization
result~\citep{akram_el-korashy_formal_2016} (not implying control-flow
correctness). He also proves soundness of control flow integrity (see above)
adapted to the capability machine, a result that seems orthogonal to the fact
that the machine uses capabilities.

%\section{Conclusion}
\label{sec:conclusion}
\todo[inline]{Insert the correct URL for the technical report in the bibtex file!}

>>>>

Capability machines originate with Dennis and Van
Horn~\cite{Dennis:1966:PSM:365230.365252} and we refer to
\citet{Levy1984capability} and \citet{Watson2015Cheri} for an overview of
previous work in the area. The capability machine formalised in
Section~\ref{sec:capab-mach-with} is a simple but representative model. It is modeled
mainly after the M-Machine~\cite{Carter:1994:HSF:195473.195579} (the enter
pointers resemble the M-Machine's) and
CHERI~\cite{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740} (the memory
capabilities resemble CHERI's). The latter is a recent and relatively mature
capability machine, which combines capabilities with a virtual memory approach,
in the interest of backwards compatibility and gradual adoption. Our local
capabilities are inspired by CHERI's, although as discussed, they can be passed
accross module boundaries, contrary to what is enforced by CHERI's default CCall
implementation.

There are plenty of other papers that enforce well-bracketed control flow at a
low level, but most are restricted to preventing particular types of attacks and
enforce only partial correctness of control flow. This includes particularly the
line of work on \emph{control-flow integrity}~\citep{abadi_control-flow_2005}.
This work uses a quite different attacker model than us: they assume an attacker
that is not able to execute code, but can overwrite arbitrary data at any time
during execution (as a model for buffer overflows). By checking the address of
every indirect jump and using memory access control to prevent overwriting code,
this work enforces what they call control-flow integrity, formalised as the
property that every jump will follow a legal path in the control-flow graph.
As discussed in Section~\ref{sec:discussion}, this correctness property does not
take into account temporal properties and seems hard to use for reasoning.

@@

Our logical relation is a unary, step-indexed Kripke logical relation (SKLR)
with recursive
worlds~\cite{Appel:2001:IMR:504709.504712,Ahmed2004semantics,Birkedal:2011:SKM:1926385.1926401},
closely related to the one used by \citet{Devriese:2016ObjCap} to formulate
capability safety in a high-level JavaScript-like lambda calculus. Our
Fundamental Theorem is similar to theirs and can be seen as a formulation of
capability safety of the capability machine. Because we are not interested in
externally observable side-effects (like console output or memory access
traces), we do not require their notion of effect parametricity. Our logical
relation uses several ideas from previous work, like Kripke worlds with regions
containing state transition systems~\citep{Ahmed:popl09}, public/private future
worlds~\citep{Dreyer:jfp12} (see Section~\ref{sec:discussion} for a discussion),
biorthogonality~\citep{pitts_operational_1998,benton_biorthogonality_2009-1,Hur:2011:KLR:1926385.1926402}.

El-Korashy also defined a formal model of a capability machine, namely CHERI,
and uses it to prove a compartmentalization
result~\citep{akram_el-korashy_formal_2016} (not implying control-flow
correctness). He also proves soundness of control-flow integrity (see above)
adapted to the capability machine, a result that seems orthogonal to the fact
that the machine uses capabilities.

%\section{Conclusion}
%\label{sec:conclusion}
\todo[inline]{Insert the correct URL for the technical report in the bibtex file!}
